---
title: 'Final Project'
subtitle: "PSTAT122: Design and Analysis of Experiments" 

author:
  - name: "Winter 2026"
affiliation-title: "Quarter"
format: 
 pdf:
    code-fold: true
    code-line-numbers: true
    code-copy: true
    code-tools: true
    self-contained: true
    toc: false
    toc-location: left
    number-sections: true
    geometry: margin=0.3in

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)

```


:::callout
<div style="text-align: center">
<span style="color: blue;"> **STUDENT NAME **</span> 
</div>
- Ethan Anghelache    (A059728)
- Miriam Montes    (4538575)
- Joanna Sells    (A591498)
- Saarah Solachuddin    (A0P1751)
:::


::: {.callout-caution}
# Due Date
**Due Date:** Monday, March 16, 2026, 11:59 PM
:::


```{r libraries, include = FALSE}
library(car)
library(emmeans)
library(tidyverse)
library(knitr)
library(broom)

```


# Introduction

The purpose of this experiment is to determine whether the visual characteristics of text influence reading speed. Specifically, we are investigating how two formatting factors, font size and color, affect the speed at which individuals read a short passage. In this experiment, we are using a $3^2$ factorial design to examine both the individual effects of font size and font color at three levels each, as well as the potential interaction between these factors. The interaction effect would indicate whether the influence of one factor changes depending on the level of the other.
 
Text formatting plays an important role in readability, especially in today’s world where we regularly read information on both digital screens and printed materials. Differences in font size and color can affect visual clarity and may influence how quickly a reader moves through a passage. Since these formatting choices are commonly used in digital content, educational resources, professional documents, and more, understanding how they influence reading speed may help inform more effective visual design decisions to improve readability.
 
Previous research has also examined how font size affects reading speed. For example, a study by Scott B. Chandler at Virginia Tech tested participants reading text in different font sizes and found that larger font sizes were generally read more quickly than smaller ones. Based on these findings, we expect that passages presented in a larger font size (18 pt) will be read more quickly than those presented in a smaller font size (10 pt). We also expect that black text may lead to faster reading speeds than red text because it is the most commonly used color in standard reading materials.


# Experimental Design

We are motivated to improve design choices for readability, thus we will be testing how visual text settings (font size, color) impact reading speed. Participants are instructed to read at the best fit speed at which information retention is highest for themselves. We define readability as the rate of reading words while maintaining a consistent level of comprehension, thus a faster rate means better readability in this experiment.

## Factors and Treatment Structure

There are two treatments, font size and font color, with three levels each: in order from highest to lowest contrast against a white background, black, red, and green font, and in order from smallest to largest text, 10 pt (small), 14 pt (medium), and 18 pt (large), creating nine unique treatment combinations:

:::callout
- Small (10 pt), green font

- Medium (14 pt), green font

- Large (18 pt), green font

- Small (10 pt), red font

- Medium (14 pt), red font

- Large (18 pt), red font

- Small (10 pt), black font

- Medium (14 pt), black font

- Large (18 pt), black font
:::

## Experimental Units

There are 16 experimental units/individual readers in total. All but six readers read three times according to their randomized treatment assignment; the six read four times. The response variable is the reading speeds measured by the time taken (in seconds) to read the same 133-word excerpt.

## Factors Classification

This experiment will be conducted as a $3^2$ factorial design with a CRD framework, with two factors and three levels per factor:

:::callout
- Factor A: Font Size. This is a fixed factor, with three levels at 10 pt, 14 pt, or 18 pt.

- Factor B: Font Color. This is a fixed factor, with three levels in green, red, or black.
:::

## Sample Size

With nine different treatment combinations, as well as six readers per treatment combination, we have a total of 54 observations/readers randomly assigned to the treatment combinations.

$$(10*3) + (6*4) = 54\text{ observations total}$$

## Design Type: Complete Randomized Design

When conducting this experiment, we did not use any blocking. With a total of 16 readers, we randomized the assignment of each treatment to three or four per person.

## Statistical Model

For this completely randomized factorial design, we have 

:::callout
1. Three levels for Factor A: Font Size

2. Three levels for Factor B: Font Color

3. Six replicates within each treatment combination
:::

Therefore, the effects model can be expressed as: $$Y_{ijk} = \mu + \tau_i + \beta_j + (\tau\beta)_{ij}+ \epsilon_{ijk}$$ with i = 1, 2, 3; j = 1, 2, 3; k = 1, 2, 3, 4, 5, 6. The error term $\epsilon$ represents the variation between people.

## Implementation

To implement randomization, a list of 16 readers was compiled in no particular order on a Google sheet, ensuring that every participant had an equal probability of being assigned to any given treatment. Each participant was randomly assigned to one of the four treatment combinations using an `R` sampling code. This ensures that every participant had an equal probability of being assigned to any given treatment.

```{r, include = FALSE}
# randomizing assignments
set.seed(21)

text_settings <- c("black, small", "black, medium", "black, large",
                   "red, small", "red, medium", "red, large",
                   "green, small", "green, medium", "green, large")

treatments <- rep(text_settings, each = 6)

sample(treatments)

```


# Data Collection

The experiment was conducted midday at each experimenter's residential home in a quiet room on March 11, 2026. Participants were recruited from the experimenters' pool of roommates. Each participant was given the same text excerpt with a specific text setting (e.g. red text color with large font size) that had been randomized and assigned per individual to control for order effects. Participants were instructed: "Read this excerpt at the usual reading speed at which you retain information well." Reading time was measured in seconds, beginning when the participant started reading and ending when they indicated they were finished.

One limitation noted during data collection was that participants were not explicitly instructed whether to read silently or aloud. While all participants appeared to follow the general instructions, those who read silently may have unintentionally skimmed the text rather than reading at their true normal pace, potentially affecting the accuracy of the recorded times. The experiment was conducted in a quiet room to minimize external distractions. No trials were repeated, as the experiment was designed as a single-attempt format. Any natural variation in performance was considered part of the data. Timing was measured consistently using a phone timer, initiated by the experimenter upon signaling "start" and stopped when the participant indicated "done."

The following tables present a portion of our collected data---categorized by our two factors and each participants' reading speed---and our key summary statistics:

```{r, echo = FALSE}
# get factors in their own columns
# font color levels (highest to lowest contrast against white background): black (-1), red (0), green (1)
# font size levels (smallest to largest font): small (-1), medium (0), large (1)
data <- read.csv("Final_Data.csv")

read_speed <- data |>
  mutate("Font Color" = factor(Text.Settings,
                               levels = c("black, small", "black, medium", "black, large",
                                          "red, small", "red, medium", "red, large",
                                          "green, small", "green, medium", "green, large"),
                               labels = rep(c("Black", "Red", "Green"), each = 3)),
         "Font Size" = factor(Text.Settings,
                              levels = c("black, small", "black, medium", "black, large",
                                          "red, small", "red, medium", "red, large",
                                          "green, small", "green, medium", "green, large"),
                              labels = rep(c("Small", "Medium", "Large"), times = 3))) |>
  rename("Time" = Time..s.) |>
  select(`Font Size`, `Font Color`, `Time`)

kable(head(read_speed, n = 9),
      caption = "Reading Time by Factors Data Collection")

```


```{r, echo = FALSE}
# get key summary stats of our data
summary_stats <- read_speed |>
  group_by(`Font Size`, `Font Color`) |>
  summarise(Mean = round(mean(Time), 2),
            SD = round(sd(Time), 2),
            N = n())

kable(summary_stats,
      col.names = c("Font Size", "Font Color", "Mean (s)", "SD (s)", "N"),
      caption = "Summary Statistics of Reading Time by Treatment Group")

```


# Analysis

## Exploratory Analysis

Before conducting formal hypothesis testing, exploratory analysis was performed to understand the structure and variability of the data. Summary statistics were computed for each of the four treatment combinations (Table 2). Reading speed seems to be quicker at a larger font, i.e. 14 pt and 18 pt, than in the smaller 10 pt font. It seems that font color has a very slight impact on the reading speed, in that when the font color is black, the average time is slightly shorter compared to its other levels per level of font size. This initial inspection suggests that font size may have a meaningful effect on reading time, while font color appears to have a comparatively minor influence.

We visualized these findings through the following figure: 

```{r, echo = FALSE, fig.align = "center"}
# boxplots to compare means
ggplot(read_speed, aes(x = `Font Size`,
                       y = `Time`,
                       fill = `Font Color`)) +
  scale_fill_manual(values = c("darkgray", "red", "green")) +
  geom_boxplot() + 
  labs(y = "Reading Time (s)",
       title = "Figure 1: Reading Time Comparisons with Boxplots") +
  theme_bw()

```

Since there are outliers present, we used the statistical outlier IQR rule to identify and remove these points from our data, resulting in the following figure:

```{r, echo = FALSE, fig.align = 'center'}
# remove outliers based on box plot
outliers_sr <- read_speed |>
  filter(`Font Size` == "Small", `Font Color` == "Red")

q1_sr <- quantile(outliers_sr$Time, 0.25)
q3_sr <- quantile(outliers_sr$Time, 0.75)
iqr_sr <- q3_sr - q1_sr

read_speed1 <- outliers_sr |>
  filter(Time <= (q1_sr - (1.5 * iqr_sr)) | Time >= (q3_sr + (1.5 * iqr_sr)))

outliers_sg <- read_speed |>
  filter(`Font Size` == "Small", `Font Color` == "Green")

q1_sg <- quantile(outliers_sg$Time, 0.25)
q3_sg <- quantile(outliers_sg$Time, 0.75)
iqr_sg <- q3_sg - q1_sg

read_speed2 <- outliers_sg |>
  filter(Time <= (q1_sg - (1.5 * iqr_sg)) | Time >= (q3_sg + (1.5 * iqr_sg)))

outliers_mg <- read_speed |>
  filter(`Font Size` == "Medium", `Font Color` == "Green")

q1_mg <- quantile(outliers_mg$Time, 0.25)
q3_mg <- quantile(outliers_mg$Time, 0.75)
iqr_mg <- q3_mg - q1_mg

read_speed3 <- outliers_mg |>
  filter(Time <= (q1_mg - (1.5 * iqr_mg)) | Time >= (q3_mg + (1.5 * iqr_mg)))

read_speed_clean <- read_speed |>
  filter(Time != read_speed1$Time,
         Time != read_speed2$Time[1],
         Time != read_speed2$Time[2],
         Time != read_speed3$Time)

# read_speed_clean

# new boxplot w/o outliers
ggplot(read_speed_clean, aes(x = `Font Size`,
                       y = `Time`,
                       fill = `Font Color`)) +
  scale_fill_manual(values = c("darkgray", "red", "green")) +
  geom_boxplot() + 
  labs(y = "Reading Time (s)",
       title = "Figure 2: Reading Time Comparisons with Boxplots No Outliers") +
  theme_bw()
```

To further investigate our pre-test analysis, we gathered observations from our new data without outliers using the following interaction plot:

```{r, echo = FALSE, fig.align = "center"}
# interaction plot
summary_stats_clean <- read_speed_clean |>
  group_by(`Font Size`, `Font Color`) |>
  summarise(Mean = round(mean(Time), 2),
            SD = round(sd(Time), 2),
            N = n())

ggplot(summary_stats_clean, aes(x = `Font Size`,
                          y = Mean,
                          color = `Font Color`,
                          group = `Font Color`)) +
  scale_color_manual(values = c("black", "red", "green")) +
  geom_point() +
  geom_line() +
  labs(y = "Mean Reading Time (s)",
       title = "Figure 3: Interaction Plot of Font Size and Font Color") + 
  theme_bw()

```

The interaction plot (Figure 3) expands on our previous observations. The three lines follow nearly the same trajectory from Small to Large font size. At the medium level, green and red font intersect, suggesting that our factors may interact and be dependent to some extent; specifically, when our font size changes from small to large, the roles of red and green font switch, and red font makes for quicker reading speed (i.e. better readability) in comparison.

Taken together, the exploratory analysis points to font size as the primary driver of a faster reading time, and font color plays a less significant role in an individual's reading speed. However, visual trends suggest there also may be an interaction between our factors. To verify these observations about our data, we will turn to hypothesis testing and a formal ANOVA test to determine any statistical significance of our observations.

## Hypothesis Testing

To conduct our hypothesis testing for this $3^2$ factorial design experiment, we looked at three sets of hypotheses:

:::callout
1. The main effect of Font Size

\begin{center}
$H_0$: $\tau_i = 0$, $\forall i$

$H_1$: $\exists$ $\tau_i \ne 0$

$\tau_i$ represents the effect of font size and is tested for changes in the reading speed results on average.
\end{center}

2. The main effect of Font Color

\begin{center}
$H_0$: $\beta_j = 0$, $\forall j$

$H_1$: $\exists$ $\beta_j \ne 0$

$\beta_j$ represents the effect of font color and is tested for changes in the reading speed results on average.
\end{center}

3. The interaction effect of Font Size and Font Color

\begin{center}
$H_0$: $(\tau \beta)_{ij} = 0$ $\forall i, j$

$H_1$: $\exists$ $(\tau \beta)_{ij} \ne 0$

$(\tau \beta)_{ij}$ represents the combination of our two treatments, and we test if this combination creates a unique joint effect that they do not have alone.
\end{center}
:::


A two-way ANOVA table tests our hypotheses simultaneously, as follows:

```{r, echo = FALSE}
# generate ANOVA table
anova <- aov(`Time` ~ `Font Size` * `Font Color`, read_speed_clean)

kable(tidy(anova),
      digits = 3,
      col.names = c("Sources of Variation", "degrees of freedom", "Sum of Squares", "Mean Square", "F", "p-value"),
      caption = "ANOVA Table")

```

From our resulting table, we can conclude that font size alone overall is the driving factor of an individual's reading speed in our data. Font size yields a p-value of 0.000, a value smaller than our significance level $\alpha = 0.05$. Thus, we reject our null hypothesis ($H_0$) that font size at all levels has equal impact on an individual's reading speed, meaning there is a statistically significant difference between the impact of small, medium, and large font size on the time it takes to read. From our previous findings, we can also conclude that our data points to larger fonts in particular as what helps an individual read faster. Thus, our data has sufficient evidence to point to larger font having better readability.

Font color as a factor yielded a p-value of 0.105, a value larger than $\alpha$, meaning we fail to reject the null hypothesis that there is no difference in impact on reading speed between the three levels, green, red, and black font. The interaction between our two factors yielded a p-value of 0.692, a value also larger than $\alpha$, indicating a failure to reject the null hypothesis, furthermore meaning there is no significant evidence of an interaction effect between the two in our data. Our two factors appear to behave independent of one other.


# Conclusions

The goal of this experiment was to determine whether font size and color of text affects reading speed. Using a $3^2$ factorial design within a completely randomized framework, we examined the effects of three font sizes (10 pt, 14 pt and 18 pt) and three font colors (black, red and green) on the time required to read a short paragraph. 

The results of the two way ANOVA indicate that font size has a statistically significant effect on reading time, whilst font color and the interaction between font size and font color are not statistically significant. With a p-value of 0.003 for font size, we reject the null hypothesis that the mean reading times are equal across all font sizes. The exploratory analysis and summary statistics show that larger font sizes tends to produce shorter reading times, suggesting that larger text improves readability and allows participants to process written information more quickly. This result is consistent with what we expected from prior research. 

In contrast, font color did not significantly influence reading speed (p=0.102). Although the descriptive statistics suggested slightly faster reading times for black text compared to red or green text, these differences were not statistically significant at the 5% level. Additionally, interaction between font size and color was not significant (p=0.939), indicating that the effect of font size on reading speed is consistent across different text colors. The interaction plot supported this conclusion, as the lines representing the three font colors remained approximately parallel across the levels of different font sizes. 

Whilst these findings provide evidence that font size is an important factor affecting reading speed, several limitations of the study should be acknowledged. The experiment included only sixteen participants, many of whom were recruited from similar social circles. This convenience sample may limit the generalizability of the results to a broader population. Additionally, participants completed multiple reading trials, which could introduce learning effects or fatigue that may influence reading performance across treatments. Participants were also not explicitly instructed whether to read silently or aloud, which may have introduced variability in reading behavior. Finally, reading times were measured manually using a phone timer, which may introduce small measurement errors. 

Future studies could strengthen the reliability of these results by using a larger and more diverse participant pool, standardizing reading instructions, and incorporating automated timing methods. Including comprehension checks or additional formatting factors such as font type, line spacing or background color could also provide further insight into how visual design influences readability. 

Overall, the results of this experiment suggest that larger font sizes improve reading speed, whilst font color alone does not significantly affect reading performance. These findings highlight the importance of font size in design considerations when presenting written information or printed material. 





# References

Comparing the Legibility and Comprehension of Type Size, Font Selection and Rendering Technology of Onscreen Type (Scott B. Chandler) https://vtechworks.lib.vt.edu/server/api/core/bitstreams/4dc7b456-3ea7-4bca-93d4-d58b53101aed/content

\newpage

# Appendices

```{r appendix, ref.label = knitr::all_labels(), echo = TRUE, eval = FALSE}
```
