---
title: 'Final Project'
subtitle: "PSTAT122: Design and Analysis of Experiments" 

author:
  - name: "Winter 2026"
affiliation-title: "Quarter"
format: 
 pdf:
    code-fold: true
    code-line-numbers: true
    code-copy: true
    code-tools: true
    self-contained: true
    toc: false
    toc-location: left
    number-sections: true
    geometry: margin=0.3in

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(error = FALSE)

```


:::callout
<div style="text-align: center">
<span style="color: blue;"> **STUDENT NAME **</span> 
</div>
- Ethan Anghelache    (A059728)
- Miriam Montes    (4538575)
- Joanna Sells    (A591498)
- Saarah Solachuddin    (A0P1751)
:::


```{r libraries, include = FALSE}
library(car)
library(emmeans)
library(tidyverse)
library(knitr)
library(broom)

```


# Introduction

The purpose of this experiment is to determine whether the visual characteristics of text influence reading speed. Specifically, we are investigating how two formatting factors, font size and color, affect the speed at which individuals read a short passage. In this experiment, we are using a $3^2$ factorial design to examine both the individual effects of font size and font color at three levels each, as well as the potential interaction between these factors. The interaction effect would indicate whether the influence of one factor changes depending on the level of the other.
 
Text formatting plays an important role in readability, especially in today’s world where we regularly read information on both digital screens and printed materials. Differences in font size and color can affect visual clarity and may influence how quickly a reader moves through a passage. Since these formatting choices are commonly used in digital content, educational resources, professional documents, and more, understanding how they influence reading speed may help inform more effective visual design decisions to improve readability.
 
Previous research has also examined how font size affects reading speed. For example, a study by Scott B. Chandler at Virginia Tech tested participants reading text in different font sizes and found that larger font sizes were generally read more quickly than smaller ones. Based on these findings, we expect that passages presented in a larger font size (18 pt) will be read more quickly than those presented in a smaller font size (10 pt). We also expect that black text may lead to faster reading speeds than red text because it is the most commonly used color in standard reading materials.


# Experimental Design

We are motivated to improve design choices for readability, thus we will be testing how visual text settings (font size, color) impact reading speed. Participants are instructed to read at the best fit speed at which information retention is highest for themselves. We define readability as the rate of reading words while maintaining a consistent level of comprehension, thus a faster rate means better readability in this experiment.

## Factor Classification, Experimental Units, & Implementation 
This experiment will be conducted as a $3^2$ factorial design with Completely Randomized Design (CRD) framework, where we have a total of 2 factors, each with 3 levels:

:::callout
- Factor A: Font Size. This is a fixed factor, with three levels at 10 pt, 14 pt, or 18 pt.

- Factor B: Font Color. This is a fixed factor, with three levels in green, red, or black.
:::
\

There are two treatments---font size and font color---with three levels each: in order from highest to lowest contrast against a white background, black, red, and green font, and in order from smallest to largest text, 10 pt, 14 pt, and 18 pt. This creates nine unique treatment combinations:

:::callout
- Small (10 pt), green font

- Medium (14 pt), green font

- Large (18 pt), green font

- Small (10 pt), red font

- Medium (14 pt), red font

- Large (18 pt), red font

- Small (10 pt), black font

- Medium (14 pt), black font

- Large (18 pt), black font
:::
\

Factor A has three levels $(a = 3)$, and Factor B has three levels $(b = 3)$. With nine different treatment combinations $(a\times b)$, we have a total of 54 observations. These observations were obtained by assigning ten readers to three treatment combinations each and six readers to four combinations each through replication. This results in a balanced treatment design where there are an equal number of replicates $(n = 6)$ for each treatment combination such that: $$N = (n\times a \times b )= 6\times3\times3 =54\text{ total observations}$$

Our experimental units are 16 individual readers spread out over 54 observations. To implement randomization, a list of our 16 readers was compiled in no particular order on a Google sheet. Each participant was randomly assigned treatment combinations using an `R` sampling code, found in the appendix. This ensures that every participant had an equal probability of being assigned to any given treatment. The response variable is the reading speeds measured by the time taken (in seconds) to read the same 133-word excerpt in three or four of the different formatting styles.

## Statistical Model

When conducting this experiment, we did not use any blocking. With a total of 16 readers, we randomized the assignment of each treatment with a code in R and replicated the list of readers and an equal number of replications for each treatment combination. For this balanced completely randomized factorial design, we have:

:::callout
1. Three levels for Factor A: Font Size

2. Three levels for Factor B: Font Color

3. Six replicates within each treatment combination
:::
\

Therefore, the effects model can be expressed as:

$$Y_{ijk} = \mu + \tau_i + \beta_j + (\tau\beta)_{ij}+ \epsilon_{ijk}$$ 
with i = 1, 2, 3; j = 1, 2, 3; k = 1, 2, 3, 4, 5, 6.

$(\tau\beta)_{ij}$ represents the interaction effect between front size and font color, and the random error term $\epsilon_{ijk}$ represents the random variation between observations, assumed to be normally distributed and i.i.d.


# Data Collection

The experiment was conducted midday at each experimenter's residential home in a quiet room on March 11, 2026. Participants were recruited from the experimenters' pool of roommates. Each participant was given the same text excerpt with a specific text setting (e.g. red text color with large font size) that had been randomized and assigned per individual to control for order effects. Participants were instructed: "Read this excerpt at the usual reading speed at which you retain information well." Reading time was measured in seconds, beginning when the participant started reading and ending when they indicated they were finished.

One limitation noted during data collection was that participants were not explicitly instructed whether to read silently or aloud. While all participants appeared to follow the general instructions, those who read silently may have unintentionally skimmed the text rather than reading at their true normal pace, potentially affecting the accuracy of the recorded times.

A notable source of variability was the difference in screen size and display properties across devices. Participants who read on a phone screen encountered smaller text and narrower line widths compared to those using a tablet or laptop, which may have influenced reading speed independent of the treatment conditions. Font rendering also differs across devices and operating systems, meaning text may have appeared sharper, larger, or more spaced out for some participants than others, all factors known to affect reading fluency.

Ambient lighting conditions were not controlled across participants. Those in brighter environments may have experienced less eye strain and read more comfortably, while participants in dimmer settings may have read more slowly due to reduced screen visibility. Similarly, although the experiment was designed to take place in a quiet room, environmental noise and other distractions could not be fully controlled for all participants, potentially disrupting concentration and inflating reading times for some individuals.

No trials were repeated, as the experiment was designed as a single-attempt format. This means any temporary lapse in focus, fatigue, or momentary distraction was captured in the data without opportunity for correction. Any natural variation in performance was considered part of the data. Timing was measured consistently using a phone timer, initiated by the experimenter upon signaling "start" and stopped when the participant indicated "done," though minor reaction time delays on either end may have introduced small inconsistencies in recorded times.

The following tables present a portion of our collected data, categorized by our two factors and each participants' reading speed and our key summary statistics:

```{r, echo = FALSE}
# get factors in their own columns
# font color levels (highest to lowest contrast against white background): black (-1), red (0), green (1)
# font size levels (smallest to largest font): small (-1), medium (0), large (1)
data <- read.csv("Final_Data.csv")

read_speed <- data |>
  mutate("Font Color" = factor(Text.Settings,
                               levels = c("black, small", "black, medium", "black, large",
                                          "red, small", "red, medium", "red, large",
                                          "green, small", "green, medium", "green, large"),
                               labels = rep(c("Black", "Red", "Green"), each = 3)),
         "Font Size" = factor(Text.Settings,
                              levels = c("black, small", "black, medium", "black, large",
                                          "red, small", "red, medium", "red, large",
                                          "green, small", "green, medium", "green, large"),
                              labels = rep(c("Small", "Medium", "Large"), times = 3))) |>
  rename("Time" = Time..s.) |>
  select(`Font Size`, `Font Color`, `Time`)

kable(head(read_speed, n = 9),
      caption = "Reading Time by Factors Data Collection")

# get key summary stats of our data
summary_stats <- read_speed |>
  group_by(`Font Size`, `Font Color`) |>
  summarise(Mean = round(mean(Time), 2),
            SD = round(sd(Time), 2))

kable(summary_stats,
      col.names = c("Font Size", "Font Color", "Mean (s)", "SD (s)"),
      caption = "Summary Statistics of Reading Time by Treatment Group (n = 6)")

```


# Analysis

As we conducted exploratory analysis, we found outliers in our data that we made the decision to remove. Following the removal of outliers and the collection of replacement observations to maintain balance, the dataset was reassembled in its final balanced form. This process is reflected below. The updated summary statistics and visualizations reflect this cleaned dataset and serve as the basis for all subsequent formal analyses.

## Exploratory Analysis

Before conducting formal hypothesis testing, exploratory analysis was performed to understand the structure and variability of the data. Summary statistics were computed for each of the four treatment combinations (Table 2). Reading speed seems to be quicker at a larger font, i.e. 14 pt and 18 pt, than in the smaller 10 pt font. It seems that font color has a very slight impact on the reading speed, in that when the font color is black, the average time is slightly shorter compared to its other levels per level of font size. This initial inspection suggests that font size may have a meaningful effect on reading time, while font color appears to have a comparatively minor influence.

We visualized these findings through the following figure: 

```{r, echo = FALSE, fig.align = "center"}
# boxplots to compare means
ggplot(read_speed, aes(x = `Font Size`,
                       y = `Time`,
                       fill = `Font Color`)) +
  scale_fill_manual(values = c("darkgray", "red", "green")) +
  geom_boxplot() + 
  labs(y = "Reading Time (s)",
       title = "Figure 1: Reading Time Comparisons with Boxplots") +
  theme_bw()

```

We visually identified four datapoints as outliers through this plot. Specifically, the minimum from treatment combination Small/Red, minimum and maximum from Small/Green, and maximum from Medium/Green were our outliers. Removing the minimum, maximum, or both from the treatment combinations identified by the boxplots as having outliers effectively uses the statistical outlier rule of 1.5 $\times$ IQR. Thus, we removed our outliers using this rule.

This, however, created an imbalance in our data. Thus, to keep our data balanced, we then produced new observations filling each treatment combination with missing values. These new observations were produced randomly and under identical experimental conditions to the initial data collection conditions.

This process resulted in an updated dataset, still balanced, but now without outliers. The following shows updated summary statistics and boxplots using this updated dataset:

```{r, echo = FALSE, fig.align = 'center'}
# remove outliers based on box plot

# outlier small/red is the min
outliers_sr <- read_speed |>
  filter(`Font Size` == "Small", `Font Color` == "Red") |>
  filter(Time == min(Time))

# outlier small/green is both the min and max
outliers_sg <- read_speed |>
  filter(`Font Size` == "Small", `Font Color` == "Green") |>
  filter(Time == max(Time) |
         Time == min(Time))

# outlier medium/green is the max
outliers_mg <- read_speed |>
  filter(`Font Size` == "Medium", `Font Color` == "Green") |>
  filter(Time == max(Time))

read_speed_clean <- read_speed |>
  filter(Time != outliers_sr$Time &
          Time != outliers_sg$Time[1] &
          Time != outliers_sg$Time[2] &
          Time != outliers_mg$Time)

read_speed_new <- read_speed_clean |>
  add_row(`Font Size` = as.factor(c("Small", "Small", "Small", "Medium")),
          `Font Color` = as.factor(c("Green", "Red", "Green", "Green")),
          Time = c(29.68, 22.88, 25.79, 22.81))

# new summary statistics
summary_stats_new <- read_speed_new |>
  group_by(`Font Size`, `Font Color`) |>
  summarise(Mean = round(mean(Time), 2),
            SD = round(sd(Time), 2))

kable(summary_stats_new,
      col.names = c("Font Size", "Font Color", "Mean (s)", "SD (s)"),
      caption = "Updated Summary Statistics of Reading Time by Treatment Group (n = 6)")

# new boxplot w/o outliers
ggplot(read_speed_new, aes(x = `Font Size`,
                             y = `Time`,
                             fill = `Font Color`)) +
  scale_fill_manual(values = c("darkgray", "red", "green")) +
  geom_boxplot() + 
  labs(y = "Reading Time (s)",
       title = "Figure 2: Updated Reading Time Comparisons with Boxplots") +
  theme_bw()

```

Our new summary statistics (Table 3) and boxplots (Figure 2) essentially lead us to our previous observations. Our data still shows reading speed as quicker at a font size larger than small, and that the impact of font color is minimal in comparison with font size.

To further investigate our pre-test analysis, we gathered observations from our new data without outliers using the following interaction plot:

```{r, echo = FALSE, fig.align = "center"}
# interaction plot
ggplot(summary_stats_new, aes(x = `Font Size`,
                          y = Mean,
                          color = `Font Color`,
                          group = `Font Color`)) +
  scale_color_manual(values = c("black", "red", "green")) +
  geom_point() +
  geom_line() +
  labs(y = "Mean Reading Time (s)",
       title = "Figure 3: Interaction Plot of Font Size and Font Color") + 
  theme_bw()

```

The interaction plot (Figure 3) expands on our previous observations. The three lines follow nearly the same trajectory from Small to Large font size. At no point do the lines intersect; they remain parallel throughout the trajectory, indicating there is no interaction present between our two factors, i.e. the factors impact reading speed independently in our data. This allows us to look at our factors' main effects independently.

Taken together, the exploratory analysis points to font size as the primary driver of a faster reading time, and font color plays a less significant role in an individual's reading speed. To verify these observations about our data, we will turn to hypothesis testing.

## Hypothesis Testing

To evaluate the effects of our factors and the interaction between them, we conducted a two-way ANOVA test at a significance level of $\alpha = 0.05$. The model to conduct this can be expressed as $$Y_{ijk} = \mu + \tau_i + \beta_j + (\tau\beta)_{ij}+ \epsilon_{ijk}$$ with i = 1, 2, 3; j = 1, 2, 3; k = 1, 2, 3, 4, 5, 6. The error term $\epsilon$ represents the variation between people.

We looked at three sets of hypotheses for this $3^2$ factorial design experiment:

:::callout
1. The main effect of Font Size

\begin{center}
$H_0$: $\tau_i = 0$, $\forall i$

There is no difference in reading speed on average between small, medium, and large font size settings.

$H_1$: $\exists$ $\tau_i \ne 0$

At least one level of font size leads to a difference in reading speed on average.

$\tau_i$ represents the effect of font size tested at all levels.
\end{center}

2. The main effect of Font Color

\begin{center}
$H_0$: $\beta_j = 0$, $\forall j$

There is no difference in reading speed on average between green, red, and black font color settings.

$H_1$: $\exists$ $\beta_j \ne 0$

At least one level of font colors leads to a difference in reading speed on average.

$\beta_j$ represents the effect of font color tested at all levels.
\end{center}

3. The interaction effect of Font Size and Font Color

\begin{center}
$H_0$: $(\tau \beta)_{ij} = 0$ $\forall i, j$

The effect of font color on reading speed does not depend on the font size.

$H_1$: $\exists$ $(\tau \beta)_{ij} \ne 0$

The effect of font color on reading speed does depend on the font size.

$(\tau \beta)_{ij}$ represents the combination of our two treatments tested at all levels to see if combinations create a unique joint effect that the factors do not have alone.
\end{center}
:::
\

Before conducting any testing, we first checked for the necessary assumptions, normality and homoscedasticity. The plots below test these assumptions:

```{r, echo = FALSE, fig.align = 'center'}
anova <- aov(`Time` ~ `Font Size` * `Font Color`, read_speed_new)

par(mfrow = c(1, 2))
plot(anova, which = 1:3)

hist(anova$residuals,
     main = "Histogram of Residuals",
     xlab = "Residuals",
     col = "lightblue")

```

From the Residuals vs Fitted plot, we see our data may be heteroscedastic, due to the slight funnel shape of the points scattered; this is further supported by the downward trend of the scale-location plot, which suggests heteroscedasticity. The Q-Q plot shows the points mostly following the line, which suggests normality; our histogram further supports this, as it follows a mostly normal shape. Thus, the ANOVA assumption for normality does hold in this case, but homoscedasticity may not. We still use the ANOVA model for analysis due to its fit to our factorial design experiment.

After checking assumptions, we began the testing. A two-way ANOVA test at a significance level of $\alpha = 0.05$ tests our hypotheses simultaneously, as follows:

```{r, echo = FALSE}
# generate ANOVA table
kable(tidy(anova),
      digits = 3,
      col.names = c("Sources of Variation", "degrees of freedom", "Sum of Squares", "Mean Square", "F", "p-value"),
      caption = "ANOVA Table")

```

From our resulting table, we can conclude that font size alone overall is the driving factor of an individual's reading speed in our data. Font size yields a p-value of 0.000, a value smaller than our significance level $\alpha = 0.05$. Thus, we reject our null hypothesis ($H_0$) that font size at all levels has equal impact on an individual's reading speed, meaning there is a statistically significant difference between the impact of small, medium, and large font size on the time it takes to read. From our previous findings, we can also conclude that our data points to larger fonts in particular as what helps an individual read faster. Thus, our data has sufficient evidence to point to larger font having better readability.

Font color as a factor yielded a p-value of 0.078, a value larger than $\alpha$, meaning we fail to reject the null hypothesis that there is no difference in impact on reading speed between the three levels, green, red, and black font. The interaction between our two factors yielded a p-value of 0.789, a value also larger than $\alpha$, indicating a failure to reject the null hypothesis, furthermore meaning there is no significant evidence of an interaction effect between the two in our data. Our two factors appear to behave independent of one other.


# Conclusions

The goal of this experiment was to determine whether font size and color of text affects reading speed. Using a $3^2$ factorial design within a completely randomized framework, we examined the effects of three font sizes (10 pt, 14 pt and 18 pt) and three font colors (black, red and green) on the time required to read a short paragraph. 

The results of the two way ANOVA indicate that font size has a statistically significant effect on reading time, whilst font color and the interaction between font size and font color are not statistically significant. With a p-value of 0.003 for font size, we reject the null hypothesis that the mean reading times are equal across all font sizes. The exploratory analysis and summary statistics show that larger font sizes tends to produce shorter reading times, suggesting that larger text improves readability and allows participants to process written information more quickly. This result is consistent with what we expected from prior research. 

In contrast, font color did not significantly influence reading speed (p=0.102). Although the descriptive statistics suggested slightly faster reading times for black text compared to red or green text, these differences were not statistically significant at the 5% level. Additionally, interaction between font size and color was not significant (p=0.939), indicating that the effect of font size on reading speed is consistent across different text colors. The interaction plot supported this conclusion, as the lines representing the three font colors remained approximately parallel across the levels of different font sizes. 

Whilst these findings provide evidence that font size is an important factor affecting reading speed, several limitations of the study should be acknowledged. The experiment included only sixteen participants, many of whom were recruited from similar social circles. This convenience sample may limit the generalizability of the results to a broader population. Additionally, participants completed multiple reading trials, which could introduce learning effects or fatigue that may influence reading performance across treatments. Participants were also not explicitly instructed whether to read silently or aloud, which may have introduced variability in reading behavior. Finally, reading times were measured manually using a phone timer, which may introduce small measurement errors. 

Future studies could strengthen the reliability of these results by using a larger and more diverse participant pool, standardizing reading instructions, and incorporating automated timing methods. Including comprehension checks or additional formatting factors such as font type, line spacing or background color could also provide further insight into how visual design influences readability. 

Overall, the results of this experiment suggest that larger font sizes improve reading speed, whilst font color alone does not significantly affect reading performance. These findings highlight the importance of font size in design considerations when presenting written information or printed material. 


# References

1. Chandler, S. B. (2001). Comparing the Legibility and Comprehension of Type Size, Font Selection and Rendering Technology of Onscreen Type. https://vtechworks.lib.vt.edu/server/api/core/bitstreams/4dc7b456-3ea7-4bca-93d4-d58b53101aed/content


\newpage
# Appendices

1. Randomization of Treatment Combinations to Assign to Subjects:

```{r, eval = FALSE}
# randomizing assignments
set.seed(21)

text_settings <- c("black, small", "black, medium", "black, large",
                   "red, small", "red, medium", "red, large",
                   "green, small", "green, medium", "green, large")

treatments <- rep(text_settings, each = 6)

sample(treatments)
```


2. Data Collection Tables (Portion of Data & Summary Statistics):

```{r, eval = FALSE}
# get factors in their own columns
# font color levels (highest to lowest contrast against white background): black (-1), red (0), green (1)
# font size levels (smallest to largest font): small (-1), medium (0), large (1)
data <- read.csv("Final_Data.csv")

read_speed <- data |>
  mutate("Font Color" = factor(Text.Settings,
                               levels = c("black, small", "black, medium", "black, large",
                                          "red, small", "red, medium", "red, large",
                                          "green, small", "green, medium", "green, large"),
                               labels = rep(c("Black", "Red", "Green"), each = 3)),
         "Font Size" = factor(Text.Settings,
                              levels = c("black, small", "black, medium", "black, large",
                                          "red, small", "red, medium", "red, large",
                                          "green, small", "green, medium", "green, large"),
                              labels = rep(c("Small", "Medium", "Large"), times = 3))) |>
  rename("Time" = Time..s.) |>
  select(`Font Size`, `Font Color`, `Time`)

kable(head(read_speed, n = 9),
      caption = "Reading Time by Factors Data Collection")

# get key summary stats of our data
summary_stats <- read_speed |>
  group_by(`Font Size`, `Font Color`) |>
  summarise(Mean = round(mean(Time), 2),
            SD = round(sd(Time), 2))

kable(summary_stats,
      col.names = c("Font Size", "Font Color", "Mean (s)", "SD (s)"),
      caption = "Summary Statistics of Reading Time by Treatment Group (n = 6)")

```


3. Boxplots of Original Data:

```{r, eval = FALSE}
# boxplots to compare means
ggplot(read_speed, aes(x = `Font Size`,
                       y = `Time`,
                       fill = `Font Color`)) +
  scale_fill_manual(values = c("darkgray", "red", "green")) +
  geom_boxplot() + 
  labs(y = "Reading Time (s)",
       title = "Figure 1: Reading Time Comparisons with Boxplots") +
  theme_bw()
```


4. Updated Data without Outliers (New Summary Statistics Table & Boxplots):

```{r, eval = FALSE}
# remove outliers based on box plot

# outlier small/red is the min
outliers_sr <- read_speed |>
  filter(`Font Size` == "Small", `Font Color` == "Red") |>
  filter(Time == min(Time))

# outlier small/green is both the min and max
outliers_sg <- read_speed |>
  filter(`Font Size` == "Small", `Font Color` == "Green") |>
  filter(Time == max(Time) |
         Time == min(Time))

# outlier medium/green is the max
outliers_mg <- read_speed |>
  filter(`Font Size` == "Medium", `Font Color` == "Green") |>
  filter(Time == max(Time))

read_speed_clean <- read_speed |>
  filter(Time != outliers_sr$Time &
          Time != outliers_sg$Time[1] &
          Time != outliers_sg$Time[2] &
          Time != outliers_mg$Time)

read_speed_new <- read_speed_clean |>
  add_row(`Font Size` = as.factor(c("Small", "Small", "Small", "Medium")),
          `Font Color` = as.factor(c("Green", "Red", "Green", "Green")),
          Time = c(29.68, 22.88, 25.79, 22.81))

# new summary statistics
summary_stats_new <- read_speed_new |>
  group_by(`Font Size`, `Font Color`) |>
  summarise(Mean = round(mean(Time), 2),
            SD = round(sd(Time), 2))

kable(summary_stats_new,
      col.names = c("Font Size", "Font Color", "Mean (s)", "SD (s)"),
      caption = "Updated Summary Statistics of Reading Time by Treatment Group (n = 6)")

# new boxplot w/o outliers
ggplot(read_speed_new, aes(x = `Font Size`,
                             y = `Time`,
                             fill = `Font Color`)) +
  scale_fill_manual(values = c("darkgray", "red", "green")) +
  geom_boxplot() + 
  labs(y = "Reading Time (s)",
       title = "Figure 2: Updated Reading Time Comparisons with Boxplots") +
  theme_bw()
```


5. Interaction Plot:

```{r, eval = FALSE}
# interaction plot
ggplot(summary_stats_new, aes(x = `Font Size`,
                          y = Mean,
                          color = `Font Color`,
                          group = `Font Color`)) +
  scale_color_manual(values = c("black", "red", "green")) +
  geom_point() +
  geom_line() +
  labs(y = "Mean Reading Time (s)",
       title = "Figure 3: Interaction Plot of Font Size and Font Color") + 
  theme_bw()
```


6. Checking Assumptions with Plots:

```{r, eval = FALSE}
anova <- aov(`Time` ~ `Font Size` * `Font Color`, read_speed_new)

par(mfrow = c(1, 2))
plot(anova, which = 1:3)

hist(anova$residuals,
     main = "Histogram of Residuals",
     xlab = "Residuals",
     col = "lightblue")
```


7. ANOVA Table:

```{r, eval = FALSE}
# generate ANOVA table
kable(tidy(anova),
      digits = 3,
      col.names = c("Sources of Variation", "degrees of freedom", "Sum of Squares", "Mean Square", "F", "p-value"),
      caption = "ANOVA Table")

```
