---
title: 'Final Project'
subtitle: "PSTAT122: Design and Analysis of Experiments" 

author:
  - name: "Winter 2026"
affiliation-title: "Quarter"
format: 
 pdf:
    code-fold: true
    code-line-numbers: true
    code-copy: true
    code-tools: true
    self-contained: true
    toc: false
    toc-location: left
    number-sections: true
    geometry: margin=0.2in

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message =  FALSE)
knitr::opts_chunk$set(warning =  FALSE)
knitr::opts_chunk$set(error =  FALSE)
knitr::opts_chunk$set(echo = FALSE)

```



:::callout
<div style="text-align: center">
<span style="color: blue;"> **STUDENT NAME **</span> 
</div>
- Ethan Anghelache    (A059728)
- Miriam Montes    (NetID 2)
- Joanna Sells    (NetID 3)
- Saarah Solachuddin    (A0P1751)
:::


::: {.callout-caution}
# Due Date
**Due Date:** Monday, March 16, 2026, 11:59 PM
:::

```{r Libraries}
library(car)
library(emmeans)
library(tidyverse)
library(knitr)
```

# Introduction
-	Clear statement of the objective or research question.
-	Brief context or motivation.

# Experimental Design
- Description of factors and treatment structure.
- Clearly state what you are measuring and the units.
Examples: Number of words recalled (count), reaction time (seconds), taste rating (1–5 scale).
- Identify which factors are fixed vs. random.
- Description of design type (CRD, RCBD, factorial, etc.).
- Explain how randomization, replication, and (if used) blocking were implemented.
- Sample size: Provide number of observations per condition.
Guideline: 5–10 per treatment for CRD, 3–5 blocks for RCBD, total feasible within 1 hour.

(You are encouraged to explore more resources for determining the sample size )

```{r}
# randomization assignments
set.seed(21)

text_settings <- c("black, small", "black, large", "red, small", "red, large")

treatments <- rep(text_settings, each = 4)

sample(treatments)
```


# Data Collection
The experiment was conducted at a residential home during midday. Participants were recruited from the experimenters' pool of roommates. Each participant was given a text excerpt that had been randomized per individual to control for order effects. Participants were instructed: "Read this excerpt at the usual reading speed at which you retain information well." Reading time was measured in seconds, beginning when the participant started reading and ending when they indicated they were finished.


One limitation noted during data collection was that participants were not explicitly instructed whether to read silently or aloud. While all participants appeared to follow the general instructions, those who read silently may have unintentionally skimmed the text rather than reading at their true normal pace, potentially affecting the accuracy of the recorded times. The experiment was conducted in a quiet room to minimize external distractions. No trials were repeated, as the experiment was designed as a single-attempt format. Any natural variation in performance was considered part of the data. Timing was measured consistently using a phone timer, initiated by the researcher upon signaling "start" and stopped when the participant indicated "done."

```{r}
data <- read.csv("Final_Data.csv")

# get factors in their own columns
# font color levels: black (-), red (+)
# font size levels: small (-), large (+)

read_speed <- data |>
  mutate(
    `Font Color` = str_detect(Text.Settings, "red, small") | str_detect(Text.Settings, "red, large"),
    `Font Size`  = str_detect(Text.Settings, "black, large") | str_detect(Text.Settings, "red, large")
  ) |>
  rename(Time = Time..s.) |>
  select(Person, `Font Color`, `Font Size`, Time)

read_speed$`Font Color` <- factor(read_speed$`Font Color`,
                                  levels = c(TRUE, FALSE),
                                  labels = c("Red", "Black"))

read_speed$`Font Size` <- factor(read_speed$`Font Size`,
                                 levels = c(TRUE, FALSE),
                                 labels = c("Large", "Small"))

kable(read_speed)
```
```{r}
summary_stats <- read_speed |>
  group_by(`Font Color`, `Font Size`) |>
  summarise(
    Mean = round(mean(Time), 2),
    SD   = round(sd(Time), 2),
    N    = n(),
    .groups = "drop"
  )

kable(summary_stats,
      col.names = c("Font Color", "Font Size", "Mean (s)", "SD (s)", "N"),
      caption   = "Table 1: Summary Statistics of Reading Time by Treatment Group",
      align     = c("l", "l", "c", "c", "c"),
      digits    = 2,
      format    = "pipe")

```
```{r}
interaction.plot(
  x.factor     = read_speed$`Font Size`,
  trace.factor  = read_speed$`Font Color`,
  response      = read_speed$Time,
  fun           = mean,
  type          = "b",
  col           = c("red", "black"),
  pch           = c(16, 17),
  lty           = 1,
  xlab          = "Font Size",
  ylab          = "Mean Reading Time (s)",
  trace.label   = "Font Color",
  main          = "Figure 1: Interaction Plot of Font Size and Font Color on Reading Time"
)


```
# Analysis
Before conducting formal hypothesis testing, exploratory analysis was performed to understand the structure and variability of the data. Summary statistics were computed for each of the four treatment combinations (Table 1). Both the Red/Large and Black/Large groups yielded mean reading times of approximately 30 seconds (30.23s and 30.53s, respectively), while the Small font groups were noticeably faster — Red/Small averaged 20.87s and Black/Small averaged 18.14s. This initial inspection suggests that font size may have a meaningful effect on reading time, while font color appears to have a comparatively minor influence.

The interaction plot (Figure 1) further supports this observation. The two lines — one for Red and one for Black — follow nearly the same downward trajectory from Large to Small font size, indicating little to no interaction between font color and font size. If a strong interaction were present, the lines would cross or diverge substantially. The near-parallel pattern suggests that the effect of font size on reading time is consistent regardless of font color. Taken together, the exploratory analysis points to font size as the primary driver of reading time, with font color playing a negligible role.

- **Hypothesis Testing**: Test your hypothesis with an appropriate statistical test (e.g., ANOVA).

- **Tables, Figures, & Code**: Include key results (ANOVA table, post-test) and relevant `R` code excerpts where needed.

-	Use `R` to analyze the data.


# Conclusions
-	Summarize key findings.
- Comment on limitations and possible improvements.


# References
 (If needed.)


# Appendices
 
 - `R` code.
